---
title: "ECON 470 Final Project"
author: "Lauren Stekler"
date: '2022-04-11'
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# Loading Data and Packages 
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
tinytex::install_tinytex() 
library(ivpack)
library(cobalt)
library(stargazer)
library(rdrobust)
library(rddensity)
library(kableExtra)
library(epiDisplay)
library(MatchIt)
library(lubridate)
library(readr)
library(hrbrthemes)
library(plotly)
library(gganimate)
library(ggthemes)
library(magick)
library(estimatr)
library(lfe)
library(rddensity)
library(here)
library(modelsummary)
library(fixest)
library(dotwhisker)
library(psych)
library(vtable)
library(ggbeeswarm)
library(ggnormalviolin)

HCRIS_Data_v1996 <- readRDS("/Users/laurenstekler/Desktop/HCRIS_Data_V1996.rds")

HCRIS_Data_v2010 <- readRDS("/Users/laurenstekler/Desktop/HCRIS_Data_v2010.rds")

HCRIS_Data <- readRDS("/Users/laurenstekler/Desktop/HCRIS_Data.rds")

HCRIS_Data_v1996_baddebt2 <- readRDS("/Users/laurenstekler/Desktop/ECON 470/HCRIS_Data_V1996_baddebt2.rds")
  
HCRIS_Data_v2010_baddebt2 <-readRDS("/Users/laurenstekler/Desktop/ECON 470/HCRIS_Data_V2010_baddebt2.rds")
  
HCRIS_Data_baddebt2 <- ("/Users/laurenstekler/Desktop/ECON 470/HCRIS_Data_baddebt2.rds")

acs_medicaid <- readRDS("/Users/laurenstekler/Desktop/acs_medicaid.rds")

insurance <- readRDS("/Users/laurenstekler/Desktop/insurance.rds")

medicaid_kff <- readRDS("/Users/laurenstekler/Desktop/medicaid-kff.rds")

```

# Preparing the Data 
```{r}
# Cleaning HCRIS and Medicaid Variable 
names(HCRIS_Data_v2010)[names(HCRIS_Data_v2010) == "state"] <- "State"

names(HCRIS_Data_v2010_baddebt2)[names(HCRIS_Data_v2010_baddebt2) == "state"] <- "State"

medicaid_kff[1, 1] <- "AL"
medicaid_kff[2, 1] <- "AK"
medicaid_kff[3, 1] <- "AZ"
medicaid_kff[4, 1] <- "AR"
medicaid_kff[5, 1] <- "CA"
medicaid_kff[6, 1] <- "CO"
medicaid_kff[7, 1] <- "CT"
medicaid_kff[8, 1] <- "DE"
medicaid_kff[9, 1] <- "DC"
medicaid_kff[10, 1] <- "FL"
medicaid_kff[11, 1] <- "GA"
medicaid_kff[12, 1] <- "HI"
medicaid_kff[13, 1] <- "ID"
medicaid_kff[14, 1] <- "IL"
medicaid_kff[15, 1] <- "IN"
medicaid_kff[16, 1] <- "IA"
medicaid_kff[17, 1] <- "KS"
medicaid_kff[18, 1] <- "KY"
medicaid_kff[19, 1] <- "LA"
medicaid_kff[20, 1] <- "ME"
medicaid_kff[21, 1] <- "MD"
medicaid_kff[22, 1] <- "MA"
medicaid_kff[23, 1] <- "MI"
medicaid_kff[24, 1] <- "MN"
medicaid_kff[25, 1] <- "MS"
medicaid_kff[26, 1] <- "MO"
medicaid_kff[27, 1] <- "MT"
medicaid_kff[28, 1] <- "NE"
medicaid_kff[29, 1] <- "NV"
medicaid_kff[30, 1] <- "NH"
medicaid_kff[31, 1] <- "NJ"
medicaid_kff[32, 1] <- "NM"
medicaid_kff[33, 1] <- "NY"
medicaid_kff[34, 1] <- "NC"
medicaid_kff[35, 1] <- "ND"
medicaid_kff[36, 1] <- "OH"
medicaid_kff[37, 1] <- "OK"
medicaid_kff[38, 1] <- "OR"
medicaid_kff[39, 1] <- "PA"
medicaid_kff[40, 1] <- "RI"
medicaid_kff[41, 1] <- "SC"
medicaid_kff[42, 1] <- "SD"
medicaid_kff[43, 1] <- "TN"
medicaid_kff[44, 1] <- "TX"
medicaid_kff[45, 1] <- "UT"
medicaid_kff[46, 1] <- "VT"
medicaid_kff[47, 1] <- "VA"
medicaid_kff[48, 1] <- "WA"
medicaid_kff[49, 1] <- "WV"
medicaid_kff[50, 1] <- "WI"
medicaid_kff[51, 1] <- "WY"

## Merging HCRIS Data and Medicaid Data
hcris.dat.merged <- HCRIS_Data_v2010 %>%
  left_join(medicaid_kff, by= "State")%>%
    mutate(expand_year = year(date_adopted), 
         expand = (year >= expand_year & !is.na(expand_year))) %>%
  rename(expand_ever = expanded)

hcris.dat.2014.merged <- hcris.dat.merged %>%
  mutate(post = (year >= 2014), treat = post*expand_ever) %>%
  filter(is.na(expand_year)|expand_year == 2014)

hcris.dat.merged <- hcris.dat.merged %>%
  mutate (post = (year >= 2014), treat = post*expand_ever)

hcris.dat.merged <- hcris.dat.merged %>%
  mutate (baddebtoperating = (bad_debt)/tot_operating_exp)

hcris.dat.2014.merged <- hcris.dat.2014.merged %>%
  mutate (baddebtoperating = (bad_debt)/tot_operating_exp)

## Merging HCRIS Data #2 (Bad Debt Variable #2)
hcris.dat.merged.baddebt2 <- HCRIS_Data_v2010_baddebt2 %>%
  left_join(medicaid_kff, by= "State")%>%
    mutate(expand_year = year(date_adopted), 
         expand = (year >= expand_year & !is.na(expand_year))) %>%
  rename(expand_ever = expanded)


hcris.dat.2014.merged.baddebt2 <- hcris.dat.merged.baddebt2 %>%
  mutate(post = (year >= 2014), treat = post*expand_ever) %>%
  filter(is.na(expand_year)|expand_year == 2014)


hcris.dat.merged.baddebt2 <- hcris.dat.merged.baddebt2 %>%
  mutate (post = (year >= 2014), treat = post*expand_ever)

hcris.dat.merged.baddebt2 <- hcris.dat.merged.baddebt2 %>%
  mutate (baddebtoperating = (bad_debt2)/tot_operating_exp)

hcris.dat.2014.merged.baddebt2 <- hcris.dat.2014.merged.baddebt2 %>%
  mutate (baddebtoperating = (bad_debt2)/tot_operating_exp)


## Cleaning Insurance Variable 
insurance[1, 1] <- "AL"
insurance[2, 1] <- "AK"
insurance[3, 1] <- "AZ"
insurance[4, 1] <- "AR"
insurance[5, 1] <- "CA"
insurance[6, 1] <- "CO"
insurance[7, 1] <- "CT"
insurance[8, 1] <- "DE"
insurance[9, 1] <- "DC"
insurance[10, 1] <- "FL"
insurance[11, 1] <- "GA"
insurance[12, 1] <- "HI"
insurance[13, 1] <- "ID"
insurance[14, 1] <- "IL"
insurance[15, 1] <- "IN"
insurance[16, 1] <- "IA"
insurance[17, 1] <- "KS"
insurance[18, 1] <- "KY"
insurance[19, 1] <- "LA"
insurance[20, 1] <- "ME"
insurance[21, 1] <- "MD"
insurance[22, 1] <- "MA"
insurance[23, 1] <- "MI"
insurance[24, 1] <- "MN"
insurance[25, 1] <- "MS"
insurance[26, 1] <- "MO"
insurance[27, 1] <- "MT"
insurance[28, 1] <- "NE"
insurance[29, 1] <- "NV"
insurance[30, 1] <- "NH"
insurance[31, 1] <- "NJ"
insurance[32, 1] <- "NM"
insurance[33, 1] <- "NY"
insurance[34, 1] <- "NC"
insurance[35, 1] <- "ND"
insurance[36, 1] <- "OH"
insurance[37, 1] <- "OK"
insurance[38, 1] <- "OR"
insurance[39, 1] <- "PA"
insurance[40, 1] <- "RI"
insurance[41, 1] <- "SC"
insurance[42, 1] <- "SD"
insurance[43, 1] <- "TN"
insurance[44, 1] <- "TX"
insurance[45, 1] <- "UT"
insurance[46, 1] <- "VT"
insurance[47, 1] <- "VA"
insurance[48, 1] <- "WA"
insurance[49, 1] <- "WV"
insurance[50, 1] <- "WI"
insurance[51, 1] <- "WY"
```


#### Part 1: Bad Debt Analysis ####

# Summary Statistics ("Bad Debt" Over Time for hospitals that expanded versus hospitals that did not expand) 
```{r}

Frac <- hcris.dat.merged%>%select(expand_ever,`baddebtoperating`)%>%na.omit()%>%group_by(expand_ever) %>% summarise_all(funs(mean, min, max), na.rm= TRUE)


colnames(Frac)= c("Expansion State", "Mean Bad Debt", "Min Bad Debt","Max Bad Debt")

kable(Frac)%>%
  kable_styling(bootstrap_options = c("striped","hover"),full_width=F)

```

# Plotting the share of bad debt over time
```{r}
hcris.dat.merged %>% group_by(year) %>% summarize(mean = mean(baddebtoperating, na.rm = TRUE)) %>% ggplot(aes(x=year, y=mean, na.rm = TRUE)) + geom_line() + geom_point() + theme_bw() + labs(
  x = "Year",
  y = "Share of Bad Debt",
  title = "Share of Bad Debt over Time"
) + 
  geom_vline(xintercept = 2013.5, color="hotpink") 


```
# Violin Plot of Bad Debt Over Time
```{r}
debt.data <- hcris.dat.merged %>%
  group_by(year) %>%
  mutate (bad_debt_low = quantile (baddebtoperating, probs = 0.01, na.rm =TRUE),
          bad_debt_high = quantile (baddebtoperating, probs = 0.99, na.rm = TRUE))%>%
  filter(baddebtoperating < bad_debt_high, baddebtoperating > bad_debt_low,
         !is.na(baddebtoperating), year>2010)


fig.debt <- debt.data %>%
  ggplot(aes(x=as.factor(year), y=baddebtoperating)) +
  geom_quasirandom(alpha=0.05, width=0.2, color = "deeppink2") +
  stat_summary(fun='median', geom='point', size=2, color = "darkblue") +
  stat_summary(fun='median', geom='line', aes(group=1), size=1.1, color = 
"#abdcf1") +
  labs(
    x="Year",
    y="Debt",
    title=""
  ) + 
  ylim (0,0.5) + 
  theme_minimal() +
  theme(
    plot.margin = unit(rep(1.2, 4), "cm"),
    plot.title = element_text(size = 19, 
                              face = "bold",
                              margin = margin(b = 8)),
    plot.subtitle = element_text(size = 14, 
                                 lineheight = 1.1,
                                 margin = margin(b = 45)),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(margin = margin(t = 5), angle = 70, hjust=1),
    axis.text.y = element_text(margin = margin(r = 5)),
    axis.title = element_text(size = 14),
    axis.title.y = element_text(margin = margin(r = 20),
                                hjust = 1),
    axis.ticks = element_line(),
    axis.title.x = element_text(margin = margin(t = 15),
                                hjust = 1),
    axis.line = element_line()
  )  

```

# Violin Plot of Bad debt in Expansion States versus Non-Expansion States
```{r}
ggplot(hcris.dat.merged, aes(x=expand_ever, y=baddebtoperating)) + geom_violin(stat="identity")

p <- ggplot(hcris.dat.merged, aes(x=expand_ever, y=baddebtoperating, fill=expand_ever)) + ylim (0, .8) + geom_violin() 
```


# Plotting bad debt over time, separately by hospitals/states that expanded Medicaid in 2014 versus those that did not

```{r}
hcris.plot.dat.merged <- hcris.dat.merged %>% filter(!is.na(expand_ever))%>%group_by(expand_ever, year)%>%summarize(mean=mean(baddebtoperating,na.rm = TRUE))

ggplot(data=hcris.plot.dat.merged, aes(x=year, y=mean, group=expand_ever, linetype = expand_ever)) + geom_line() + geom_point()+theme_bw() + geom_vline(xintercept = 2013.5, color = "hotpink")+
  geom_text(data = hcris.plot.dat.merged %>% filter(year == 2016), aes(label = c("Non-Expansion States","Expansion States"),
                                                              x = year + 1,
                                                              y = mean)) +
  guides(linetype="none")  +
  labs(
    x= "Year",
    y = "Portion of Uninsured",
    Title = "Portion of Uninsured Over Time")

```


# Estimating the effect of Medicaid expansion on bad debt using a difference-in-differences regression estimator. Only  focusing  on hospitals/states that expanded in 2014 versus those that never expanded. 
```{r}

summary(lm(baddebtoperating ~ post + expand_ever + treat, data=hcris.dat.2014.merged))

```

# Including hospital and year fixed effects in the estimates. 
```{r}
summary(felm(baddebtoperating ~ treat | factor (provider_number) + factor(year), data = hcris.dat.2014.merged))

```

# Repeating the analysis above, but including all states (even those that expanded after 2014).
```{r}

summary(felm(baddebtoperating ~ treat | factor (provider_number) + factor(year), data = hcris.dat.merged))

```

#  Creating an “event study” graph showing the effects of Medicaid expansion in each year. Only for hospitals/states that expanded in 2014 versus hospitals/states that never expanded. 
```{r}

event.dat <- hcris.dat.2014.merged %>%
  mutate(expand_2012 = expand_ever*(year == 2012), 
         expand_2013 = expand_ever*(year == 2013),
         expand_2014 = expand_ever*(year == 2014), 
         expand_2015 = expand_ever*(year == 2015),
         expand_2016 = expand_ever*(year == 2016), 
         expand_2017 = expand_ever*(year == 2017))

event.hcris.reg <- lm(baddebtoperating ~ expand_2012 + expand_2014 + expand_2015 + expand_2016 + expand_2017 + factor(year) + factor(provider_number), data=event.dat)

point.est <- as_tibble(c(event.hcris.reg$coefficients[c("expand_2012","expand_2014","expand_2015",
                                            "expand_2016","expand_2017")]),
                       rownames = "term")

ci.est <- as_tibble(confint(event.hcris.reg)[c("expand_2012","expand_2014","expand_2015",
                                   "expand_2016","expand_2017"),],
                    rownames = "term")

point.est <- point.est %>% rename(estimate = value)

ci.est <- ci.est %>% rename(conf.low = `2.5 %`, conf.high = `97.5 %`)

new.row <- tibble(
  term = "expand_2013",
  estimate = 0,
  conf.low = 0,
  conf.high = 0,
  year = 2013
) 

event.plot.dat<-point.est%>%
  left_join(ci.est, by=c("term"))%>%
  mutate(year = c(2012, 2014, 2015, 2016, 2017))%>%
  bind_rows(new.row)%>%
  arrange(year)

dwplot(event.plot.dat,
       vline=geom_vline(xintercept = 0, linetype = 2), 
       vars_order = c("expand_2018","expand_2017","expand_2016","expand_2015","expand_2014","expand_2013","expand_2012"),
       whisker_args = list(color="black", size = 1.1),
       dot_args = list(color= "black"))+
  coord_flip() + theme_bw() +theme(legend.position = "none") +
                         labs(y = "Year",
                              x = "Estimate and 95% CI",
                              title = "Effects of Medicaid Expansion on Bad Debt in Each Year") + 
                         scale_y_discrete(labels = c ("expand_2012" = "2012",
                                                      "expand_2013" = "2013",
                                                      "expand_2014" = "2014",
                                                      "expand_2015" = "2015",
                                                      "expand_2016" = "2016",
                                                      "expand_2017" = "2017"))


```
#### Part 2: Analyzing Heterogeneity: Bed Size ####
```{r}
# Creating a scatterplot
ggplot(hcris.dat.merged, aes(x=baddebtoperating, y=beds, color=expand_ever), na.rm = TRUE) + 
    geom_point(size=0.5) +
  ylim(0, 2000) +
  xlim(0, 1.25) + 
    theme_ipsum()  


ggplot(hcris.dat.merged, aes(x=baddebtoperating, y=beds, color=expand_ever), na.rm = TRUE) + 
    geom_point(size=0.5) +
  ylim(0, 2000) +
  xlim(0, 1.25) + 
    theme_ipsum() 


```

#### Part 4: Analyzing Heterogeneity: Unique Bad Debt Measurements ####

# Summary Statistics ("Bad Debt" Over Time for hospitals that expanded versus hospitals that did not expand) 
```{r}

Fral <- hcris.dat.merged.baddebt2%>%select(expand_ever,`baddebtoperating`)%>%na.omit()%>%group_by(expand_ever) %>% summarise_all(funs(mean, min, median, max), na.rm= TRUE)

colnames(Fral)= c("Expansion State", "Mean Bad Debt", "Min Bad Debt","Median Bad Debt","Max Bad Debt")

kable(Fral)%>%
  kable_styling(bootstrap_options = c("striped","hover"),full_width=F)

```

# Plotting the share of bad debt over time
```{r}
hcris.dat.merged.baddebt2 %>% group_by(year) %>% summarize(mean = mean(baddebtoperating, na.rm = TRUE)) %>% ggplot(aes(x=year, y=mean, na.rm = TRUE)) + geom_line() + geom_point() + theme_bw() + labs(
  x = "Year",
  y = "Share of Bad Debt",
  title = "Share of Bad Debt over Time"
) + 
  geom_vline(xintercept = 2013.5, color="hotpink") 

```


# Plotting bad debt over time, separately by hospitals/states that expanded Medicaid in 2014 versus those that did not
```{r}
hcris.plot.dat.merged.baddebt2 <- hcris.dat.merged.baddebt2 %>% filter(!is.na(expand_ever))%>%group_by(expand_ever, year)%>%summarize(mean=mean(baddebtoperating,na.rm = TRUE))

ggplot(data=hcris.plot.dat.merged.baddebt2, aes(x=year, y=mean, group=expand_ever, linetype = expand_ever)) + geom_line() + geom_point()+theme_bw() + geom_vline(xintercept = 2013.5, color = "hotpink")+
  geom_text(data = hcris.plot.dat.merged.baddebt2 %>% filter(year == 2016), aes(label = c("Non-Expansion States","Expansion States"),
                                                              x = year + 1,
                                                              y = mean)) +
  guides(linetype="none")  +
  labs(
    x= "Year",
    y = "Portion of Uninsured",
    Title = "Portion of Uninsured Over Time")

```
# Estimating the effect of Medicaid expansion on bad debt using a difference-in-differenecs regression estimator. Only focusing on hospitals/states that expanded in 2014 versus those that never expanded.
```{r}

summary(lm(baddebtoperating ~ post + expand_ever + treat, data=hcris.dat.2014.merged.baddebt2))

```

# Including hospital and year fixed effects in the regerssion above.  
```{r}
summary(felm(baddebtoperating ~ treat | factor (provider_number) + factor(year), data = hcris.dat.2014.merged.baddebt2))

```

# Repeat the analysis above, but including all states (even those that expanded after 2014).
```{r}

summary(felm(baddebtoperating ~ treat | factor (provider_number) + factor(year), data = hcris.dat.merged.baddebt2))

```
